---
title: "sodalake-sip Calisp vs IRMS"
output:
  html_document:
    df_print: paged
---

R markdown using calisp.py output. Importing and wrangling of feather data adapted form MS python jupyter notebook.
Here we want to compare the bulk isotope ratio change per sample between calculated from total protein (Calis-p) and measured from whole biomass (IRMS)

Difference from the old R script: bypass bin level, directly get the sample isotope ratio from the median of peptides isotope ratio

# Load libraries and basic settings

Setup working directory
```{r setup, include=FALSE}
# setup chuck to make working directory set for all chunks
wd = "/yihualiu/mscalisp/"
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = wd)
```

Load libraries

```{r}
library("stringr")
library("ggplot2")
library("tidyr")
library("dplyr")
library("readr")
library("RColorBrewer")
library("tidytable")
library("Hmisc")
library("svglite")
```

## Input Parameters
First choice from N (15N) and C (13C)

The input wrangeled files for 13C are in "./GEL_C/wrangeled_output", for 15N are in "./GEL_N/wrangeled_output"

Also specify other parameters
```{r}
la = "C"                                              # specify the isotope label, "C" for 13C and "N" for 15N # nolint
# la = "N"

sodataxo_dir = str_c(wd,"Soda lake MAGs short taxonomy.csv")   # directory of reference file with bin taxonomic annotation
irmsfile = str_c(wd,"2019_IRMS.csv")                  # specify the IRMS file name/directory
binls = c("bin25_GEM2", "bin31_GEM2", "bin17_LCM1")   # specify the bins to address when visualizing peptide isotope ratio distribution

```

Generate the other variables depends on C or N
```{r}
# wrangeled calisp data dir
  wrangeled_data_dir = str_c("GEL_",la,"/wrangled_output/")
# output dir
outroot = str_c("GEL_",la,"/analysis_", Sys.Date(),"/")
  dir.create(outroot)
outIRMS = str_c(outroot, "IRMS_Calisp/")
  dir.create(outIRMS)
```

## Import dataset

```{r}
data_all = read_csv(str_c(wrangeled_data_dir,"calisp_out_good_peptides.csv"))
data_unique = read_csv(str_c(wrangeled_data_dir,"calisp_out_unique_peptides.csv"))
```

# Calculate the protein-based isotope ratio of sample
## Sample peptide isotope distribution
First, have a look at the isotope ratio distribution of peptides from each sample.

This is similar to GEL_sip_anal_2, but here we plot at level of sample, and explore the impact of two approaches 
in GEL_sip_anal_1 to assign peptides that mapped to a cluster of multiple proteins and bins.

Here are the two ways:
  1. DIscard all peptides that mapped to more than one proteins or bins in a sample (data_unique)
  2. Keep the peptides that maps to multiple bins, since this does not impact the bulk analysis on sample level (data_all)

We use the same output directory as in GEL_sip_anal_2, and create the directory if it does not exist.
```{r}
# create a directory for peptide distribution
boxdir = str_c(outroot, "/peptide_distribution/")
dir.create(boxdir)
```

### Functions to process data and create plot boxplots
```{r}
weighted_box_stats = function(data){
  data%>%
  select(sample_ID, exp, hour, control, pattern_total_intensity, ratio_na) %>%
  # Assign sample_ID "" to 
  group_by(sample_ID, exp, hour, control) %>%
  summarise(minRo = min(ratio_na),
            y25 = wtd.quantile(ratio_na, weights = pattern_total_intensity, probs = 0.25),
            y50 = wtd.quantile(ratio_na, weights = pattern_total_intensity, probs = 0.5),
            y75 = wtd.quantile(ratio_na, weights = pattern_total_intensity, probs = 0.75),
            maxRo = max(ratio_na)) %>%
  # Note: in geom_boxplot, ymin is not the min number of y, but  
  #  "smallest observation greater than or equal to lower hinge (q1) - 1.5 * IQR"
  # where IQR ( inter-quartile range) = q3-q1
  # here instead, the following code use min of y or q1 - 1.5*IQR, whichever greater
  # similarly, ymax = max of y or q3 + 1.5*IQR, whichever smaller
  mutate(yIQR = y75 - y25) %>%
  mutate(y0 = ifelse(y25 - 1.5 * yIQR > minRo, y25 - 1.5 * yIQR, minRo),
         y100 = ifelse(y75 + 1.5 * yIQR < maxRo, y75 + 1.5 * yIQR, maxRo),
         hour = as.numeric(hour)) %>% 
  mutate(exp = ifelse(sample_ID == "TI1", "Initial_1", 
                  ifelse(sample_ID == "TI2", "Initial_2", exp)))
  # Note: I found there is a ploting bug if I merge the above two mutate() into one, but I don't know why
}

unweighted_box_stats = function(data){
  data %>%
  select(sample_ID, exp, hour, control, pattern_total_intensity, ratio_na) %>%
  group_by(sample_ID, exp, hour, control) %>%
  summarise(minRo = min(ratio_na),
            y25 = wtd.quantile(ratio_na, probs = 0.25),
            y50 = wtd.quantile(ratio_na, probs = 0.5),
            y75 = wtd.quantile(ratio_na, probs = 0.75),
            maxRo = max(ratio_na)) %>%
  # Note: in geom_boxplot, ymin is not the min number of y, but  
  #  "smallest observation greater than or equal to lower hinge (q1) - 1.5 * IQR"
  # where IQR ( inter-quartile range) = q3-q1
  # here instead, the following code use min of y or q1 - 1.5*IQR, whichever greater
  # similarly, ymax = max of y or q3 + 1.5*IQR, whichever smaller
  mutate(yIQR = y75 - y25) %>%
  mutate(y0 = ifelse(y25 - 1.5 * yIQR > minRo, y25 - 1.5 * yIQR, minRo),
         y100 = ifelse(y75 + 1.5 * yIQR < maxRo, y75 + 1.5 * yIQR, maxRo),
         hour = as.numeric(hour)) %>% 
  mutate(exp = ifelse(sample_ID == "TI1", "Initial_1", 
                  ifelse(sample_ID == "TI2", "Initial_2", exp)))
}


# function to plot manually calculated stats
box_manual = function(stats, weight ="Un", uniquepep = 0, title_prfx =""){
# testing data
# stats = t1 %>%
#     mutate(hour = as.numeric(hour))%>%
#   mutate(exp = ifelse(sample_ID == "TI1", "Initial_1", 
#                   ifelse(sample_ID == "TI2", "Initial_2", exp)))
# weight = "Un"
# uniquepep = 0

  # sample info shown in plot title
  sampleinfo = ifelse(uniquepep, "from all unique peptides", "from all good quality peptides")

  # calculate the blank quantiles' average over time, as reference lines to show on the diagram
  blk <- stats %>%
    filter(control == TRUE) 

  y25ave = mean(blk$y25)
  y50ave = mean(blk$y50)
  y75ave = mean(blk$y75)

  # plot
  if (la == "C") {
    stats %>% 
    # t1 %>%
    # mutate(hour = as.numeric(hour))%>%
    # mutate(exp = ifelse(sample_ID == "TI1", "Initial_1", 
    #               ifelse(sample_ID == "TI2", "Initial_2", exp))) %>%
    # filter(exp != "Initial_1") %>%
    ggplot(., aes(hour, group = hour)) +
    geom_boxplot(aes(x= hour, 
                     ymin = y0, 
                     lower = y25, middle = y50, upper = y75, ymax = y100, 
                     fill = exp), alpha = 0.8,
                 stat = "identity")+
    facet_grid(~exp) +

    scale_fill_manual(values = c("Blank" = "#59a9eb", "Carbon" = "#25c450"))+
    ylim(0.5, 1.5)+
    ylab( str_c("Peptide 13C/12C (%) per sample"))+
    theme_bw()+
    #Add lines for the mean y75, mean y75 and mean y50 of blank, values were calculated externally
    geom_hline(aes(yintercept = y75ave), linetype = 1, color = "#949494")+
    geom_hline(aes(yintercept = y25ave), linetype = 1, color = "#949494")+
    geom_hline(aes(yintercept = y50ave), linetype = 2, color = "#949494")+
    labs(title = str_c(title_prfx,weight,"weighted C isotope distribution, ",sampleinfo))
  } else {
   stats %>% 
      ggplot(., aes(hour, group = hour)) +
      geom_boxplot(aes(x= hour, 
                       ymin = y0, 
                       lower = y25, middle = y50, upper = y75, ymax = y100, 
                       fill = exp), alpha = 0.8,
                   stat = "identity")+
      facet_grid(~exp)+
      scale_fill_manual(values = c("Blank" = "#59a9eb", "Carbon" = "#25c450", "Ammo" = "#c2083c", "Urea" = "#ffcd00", "N2" = "#e0c4ad"))+
      ylim(-1.5, 2)+
      ylab( str_c(" Protein 15N/14N (%) per sample"))+
      theme_bw()+
      #Add lines for the mean y75, mean y75 and mean y50 of blank, values were calculated externally
      geom_hline(aes(yintercept = y75ave), linetype = 1, color = "#949494")+
      geom_hline(aes(yintercept = y25ave), linetype = 1, color = "#949494")+
      geom_hline(aes(yintercept = y50ave), linetype = 2, color = "#949494")+
    labs(title = str_c(title_prfx, weight,"weighted N isotope distribution, ",sampleinfo))
  }
}

# function to save the plots
box_save = function(MAG="", weight ="Un", uniquepep = 0){
  # sample info shown in saved plot file name
  sampleinfo_filename = ifelse(uniquepep, "unique_pep", "all_pep")
  
  s = ifelse(MAG == "","","_")
  
   ggsave(str_c(boxdir,MAG,s,la,"_",weight,"weighted_na_ratio_",sampleinfo_filename,"_2019GEL.svg"), plot= last_plot(), device = "svg", width = 8, height = 8.5, units = "in")
   ggsave(str_c(boxdir,MAG,s,la,"_",weight,"weighted_na_ratio_",sampleinfo_filename,"_2019GEL.png"), plot= last_plot(), device = "png", width = 8, height = 8.5, units = "in")
}

```

### Plotting
All good quality protein, unweighted
```{r}
data_all %>%
  unweighted_box_stats() %>%
  box_manual()
```

```{r}
box_save()
```

All good quality peptides, pattern_total_intensity weighted
```{r}
data_all %>%
  weighted_box_stats() %>%
  box_manual(., "Pattern intensity ")
```

```{r}
box_save(weight = "")
```

Only peptides that maps to one protein (unique peptides), unweighted
```{r}
data_unique %>%
  unweighted_box_stats() %>%
  box_manual(., ,uniquepep = 1)
```

```{r}
box_save(uniquepep = 1)
```
Only peptides that maps to one protein (unique peptides), pattern_total_intensity weighted
```{r}
data_unique %>%
  weighted_box_stats() %>%
  box_manual(., "Pattern intensity ",uniquepep = 1)
```

```{r}
box_save(weight = "", uniquepep = 1)
```

Loop to plot for selected MAGs

```{r}
for (bin in binls) {
  df_plot = data_unique %>%
    filter(bins == bin) 
  # Unweighted
  df_plot %>%
    unweighted_box_stats() %>%
    box_manual(., ,uniquepep = 1, title_prfx = str_c(bin, " "))  
  
  box_save(MAG=bin, weight = "Un", uniquepep = 1)
  
  # Pattern intensity Weighted
  df_plot %>%
    weighted_box_stats() %>%
    box_manual(., "Pattern intensity ",uniquepep = 1, title_prfx = str_c(bin, " "))
  
  box_save(MAG=bin, weight = "", uniquepep = 1)
}
```

# Sample IRMS isotope ratio
load IRMS measured C isotope ratio for each sample

```{r}
# read IRMS data
irms <- read_csv(irmsfile) %>%
  replace(.=="TI","0") %>%
  replace(.=="Blk", "Blank") %>%
  replace(.=="Uni", "Blank")%>%
  replace(.=="Cab", "Carbon")%>%
  replace(.=="Nit", "N2")%>%
  replace(.=="Ure", "Urea") %>%
  replace(.=="Amo", "Ammo") %>%
  # 20200412
  # Convert IRMS delta to ratio, note std used came from Calis-p but not IRMS
  # C VPDB standard 13C/12C = 0.0111802 in Isodat software
  # 15N/14N = 0.003676433/0.996323567
  mutate(Ratio_13C = (d13C_IRMS/1000 + 1)* 0.0111802,
         exp = Label,
         Ratio_15N = (d15N_IRMS/1000 + 1)* 0.003676433/0.996323567,
         Time = as.integer(Time)) %>%
  select(-Label)

#only carbon or nitrogen samples
if(la == "C"){
  irms_la <- irms %>%
    filter(exp %in% c("Blank", "Carbon")) %>%
    select(-c(d15N_IRMS, Npct, Ratio_15N)) %>%
    rename(time = Time, hours = Hours, rep = Rep, delta_irms = d13C_IRMS, ratio_irms = Ratio_13C)
} else {
  irms_la <- irms %>%
    filter(exp %in% c("Blank", "N2", "Urea", "Ammo")) %>%
    select(-c(d13C_IRMS, Cpct, Ratio_13C)) %>%
  rename(time = Time, hours = Hours, rep = Rep, delta_irms = d15N_IRMS, ratio_irms = Ratio_15N)
}
```


## Sample peptide vs IRMS isotope ratio
Calculate the isotope ratio of sample from the median of the isotope ratio of peptides
From:
  1. Same filtering criteria as the one used to plot peptide ratio (data_unique)
  2. data_unique + peptides that maps to multiple bins, since this does not impact the bulk analysis on sample level (data_all)

Prepare library and functions
```{r}
# limma library for weighted median
library(limma)

# function to calculate median sample calisp na ratio and join with irms
ra_join = function(calispdf, irmsdf){
  calispdf %>%
      # get median peptide ratio as sample ratio. Use median and pattern_total_intensity weighted median
    summarise(ratio_na = median(ratio_na), ratio_na.weight = weighted.median(ratio_na, pattern_total_intensity),.by = c(sample_ID, control)) %>%
    # join irms and Calisp results
    # Also assign t0 blank calisp value to TI samples (use the "Sample_ID" from irms)
    full_join(., irmsdf, by = "sample_ID") %>%
    # assign "TI" to the time of TI1 and TI2
    mutate(time = as.character(time))%>%
    mutate(time = ifelse(sample_ID %in% c("TI1","TI2"),"TI",time))
}


```
Get data from files and prepare table to plot
```{r}
# reload data from file
data_all = read_csv(str_c(wrangeled_data_dir,"calisp_out_good_peptides.csv"))
data_unique = read_csv(str_c(wrangeled_data_dir,"calisp_out_unique_peptides.csv"))

# from data_all
data_calispall_irms = ra_join(data_all, irms_la) %>%
  mutate(uniq_pep = FALSE)
# from data_unique
data_calispunique_irms = ra_join(data_unique, irms_la) %>%
  mutate(uniq_pep = TRUE)

# combine to make one table to report
data_calisp_irms = bind_rows(data_calispall_irms, data_calispunique_irms)

write_csv(data_calisp_irms, str_c(outIRMS, "Calisp-IRMS.csv"))

```

Discard raw data to free up memory
```{r}
remove(data_raw, irms, irms_la, data_all, data_unique, data_calispall_irms, data_calispunique_irms)
gc()
```


## Plot IRMS against Calis-p results
Here we want to compare the bulk isotope ratio change per sample between calculated from total protein (Calis-p) and measured from whole biomass (IRMS)

## Load data
```{r}
data_calisp_irms = read_csv(str_c(outIRMS, "Calisp-IRMS.csv"))
```

## 1. Unique peptides vs all peptides
Take a look on how big the difference is, when calculating sample calisp na ratio from all peptides or just unique peptides  
Create prefix for axis label depends on N or C
```{r}
if (la == 'C') {
  a <- 13
  b <- 12
  # pref = expression(""^{13}*"C/"^{12}*"C ")
} else {
  a <- 15
  b <- 14
  # pref = expression(""^{15}*"N/"^{14}*"N ")
}

```

```{r}
data_calisp_irms %>%
  ggplot()+
  geom_point(aes(x= ratio_na, y = ratio_irms, color = uniq_pep), shape = 16,  size =5.4, alpha = 0.8) +
  geom_text(aes(x= ratio_na, y = ratio_irms,label=time), color="White", size = 3.8)+
  scale_color_discrete(breaks=c(TRUE,FALSE),
                      labels=c("Unique peptides", "All good peptides"))+
  facet_wrap(~exp, scales = "fixed")+
  xlab(substitute("Protein "^{a}*la*"/"^{b}*la*" (Calis-p Protein-SIP)", list(a = a, b = b, la = la))) +
  ylab(substitute("Total Biomass "^{a}*la*"/"^{b}*la*" (CF-EA-IRMS)", list(a = a, b = b, la = la)))+
  theme_bw()+
  theme(panel.grid.minor = element_blank(), legend.title = element_blank(), legend.position = "bottom")+
  labs(title = substitute(""^{a}*la*"/"^{b}*la*" of bulk biomass and protein", list(a = a, b = b, la = la)))
```
Save the plot
```{r}
ggsave(str_c(outIRMS, "Calisp-IRMS_UNIvsALLpep.png"), width = 10, height = 5)
ggsave(str_c(outIRMS, "Calisp-IRMS_UNIvsALLpep.svg"), device = "svg",width = 10, height = 5)
```

Show hours as color
```{r}
data_calisp_irms %>%
  ggplot()+
  geom_point(aes(x= ratio_na, y = ratio_irms, color = hours, shape = uniq_pep),  size =5.4, alpha = 0.6) +
  geom_text(aes(x= ratio_na, y = ratio_irms,label=time), color="White", size = 3.8)+
  scale_color_gradient(low="blue", high = "red", name = "Incubation\nHours")+
  scale_shape_manual(breaks=c(TRUE,FALSE),
                     name = substitute("Protein "^{a}*la*"/"^{b}*la*" From:", list(a = a, b = b, la = la)),
                     values=c(15, 16),
                     labels = c("Unique peptides", "All good peptides"))+
  facet_wrap(~exp, scales = "fixed")+
  xlab(substitute("Protein "^{13}*"C/"^{12}*"C (Calis-p Protein-SIP)", list(a = a, b = b, la = la)))+
  ylab(substitute("Total Biomass "^{13}*"C/"^{12}*"C (CF-EA-IRMS)", list(a = a, b = b, la = la)))+
  theme_bw()+
  theme(panel.grid.minor = element_blank(), legend.position = "bottom")+
  labs(title = substitute(""^{a}*la*"/"^{b}*la*" of bulk biomass and protein", list(a = a, b = b, la = la)))

```
```{r}
# save the file
ggsave(str_c(outIRMS,"IRMS-Calisp_UNIvsALLpep_hour-colored",".png"), plot = last_plot(), device = "png", width = 10, height = 6)
ggsave(str_c(outIRMS,"IRMS-Calisp_UNIvsALLpep_hour-colored",".svg"), plot = last_plot(), device = "svg", width = 10, height = 6)

```

How about showing blank and carbon in one plot
```{r}
data_calisp_irms %>%
  ggplot()+
  geom_point(aes(x= ratio_na, y = ratio_irms, color = uniq_pep, shape = exp),  size =5.4, alpha = 0.8) +
  geom_text(aes(x= ratio_na, y = ratio_irms,label=time), color="White", size = 3.8) +
  xlab(substitute("Protein "^{a}*la*"/"^{b}*la*" (Calis-p Protein-SIP)", list(a = a, b = b, la = la)))+
  ylab(substitute("Total Biomass "^{a}*la*"/"^{b}*la*" (CF-EA-IRMS)", list(a = a, b = b, la = la)))+
  scale_shape_discrete(name="")+
  scale_color_discrete(breaks=c(TRUE,FALSE),
                       name = substitute("Protein "^{a}*la*"/"^{b}*la*" From:", list(a = a, b = b, la = la)),
                      labels=c("Unique peptides", "All good quality peptides"))+
  theme_bw()+
  theme(panel.grid.minor = element_blank(), legend.position = "bottom")+
  labs(title = substitute(""^{a}*la*"/"^{b}*la*" of bulk biomass and protein", list(a = a, b = b, la = la)))

```
```{r}
# save the file
ggsave(str_c(outIRMS,"IRMS-Calisp_UNIvsALLpep_NOfecet",".png"), plot = last_plot(), device = "png", width = 10, height = 6)
ggsave(str_c(outIRMS,"IRMS-Calisp_UNIvsALLpep_NOfecet",".svg"), plot = last_plot(), device = "svg", width = 10, height = 6)
```

Now plot ratio_irms against hour, with Calisp ratio (ratio_na) from all good peptides as color gradient, fecet by exp
```{r}
data_calisp_irms %>%
  filter(uniq_pep == FALSE) %>%
  ggplot()+
  geom_point(aes(x= hours, y = ratio_irms*100, color = ratio_na),  size =5.4, alpha = 0.8) +
  geom_text(aes(x= hours, y = ratio_irms*100,label=time), color="White", size = 3.8) +
  xlab("Incubation Hours")+
  ylab(substitute("Total Biomass ("^{a}*la*"/"^{b}*la*")% (CF-EA-IRMS)", list(a = a, b = b, la = la)))+
  scale_shape_discrete(name="")+
  scale_color_gradient(low= "#5063f2", high = "#f7743b", name = substitute("Protein ("^{a}*la*"/"^{b}*la*")% (Calis-p Protein-SIP)", list(a = a, b = b, la = la)))+
  facet_wrap(~exp, scales = "fixed")+
  theme_bw()+
  theme(panel.grid.minor = element_blank(), legend.position = "bottom")#+
  # labs(title = substitute(""^{a}*la*"/"^{b}*la*" of bulk biomass and protein", list(a = a, b = b, la = la)))
```
```{r}
# save the file
ggsave(str_c(outIRMS,"IRMS-hour_allpep_Calisp-colored_fecet",".png"), plot = last_plot(), device = "png", width = 6, height = 4)
ggsave(str_c(outIRMS,"IRMS-hour_allpep_Calisp-colored_fecet",".svg"), plot = last_plot(), device = "svg", width = 6, height = 4)
```

Now overlay the fecets, with grey color for "Blank" samples and color gradient for experiment samples

```{r}
ggplot()+
  geom_point(data = data_calisp_irms %>% filter((uniq_pep == FALSE)&(exp == "Blank")), aes(x= hours, y = ratio_irms), color = "#4d4d4d",  size =5.4, alpha = 0.8)+
  geom_point(data = data_calisp_irms %>% filter((uniq_pep == FALSE)&(exp != "Blank")), aes(x= hours, y = ratio_irms, color = ratio_na),  size =5.4, alpha = 0.8) +
  scale_color_gradient(low= "#5063f2", high = "#f7743b", name = substitute("Protein "^{a}*la*"/"^{b}*la*" (Calis-p Protein-SIP), Blank is in black", list(a = a, b = b, la = la))) +
  # geom_text(data = data_calisp_irms %>% filter((uniq_pep == FALSE)&(exp == "Blank")), aes(x= hours, y = ratio_irms,label=ratio_na), color="White", size = 3.8) +
  # geom_text(data = data_calisp_irms %>% filter((uniq_pep == FALSE)&(exp == "Carbon")), aes(x= hours, y = ratio_irms,label=ratio_na), color="White", size = 3.8) + 
  xlab("Incubation Hours")+
  ylab(substitute("Total Biomass "^{a}*la*"/"^{b}*la*" (CF-EA-IRMS)", list(a = a, b = b, la = la)))+
  theme_bw()+
  theme(panel.grid.minor = element_blank(), legend.position = "bottom")+
  labs(title = substitute(""^{a}*la*"/"^{b}*la*" of bulk biomass and protein", list(a = a, b = b, la = la)))
```
```{r}
# save the file
ggsave(str_c(outIRMS,"IRMS-hour_allpep_Calisp-colored_greyblk",".png"), plot = last_plot(), device = "png", width = 10, height = 6)
ggsave(str_c(outIRMS,"IRMS-hour_allpep_Calisp-colored_greyblk",".svg"), plot = last_plot(), device = "svg", width = 10, height = 6)
```

With color for both blank and labeled
```{r}
ggplot() +
  geom_point(data = data_calisp_irms %>% filter(uniq_pep == FALSE), aes(x= hours, y = ratio_irms, color = ratio_na, shape = exp),  size =5.4, alpha = 0.8) +
  scale_color_gradient(low= "#5063f2", high = "#f7743b", name = substitute("Protein "^{a}*la*"/"^{b}*la*" (Calis-p Protein-SIP)", list(a = a, b = b, la = la))) +
  scale_shape_discrete(name="")+
  xlab("Incubation Hours")+
  ylab(substitute("Total Biomass "^{a}*la*"/"^{b}*la*" (CF-EA-IRMS)", list(a = a, b = b, la = la)))+
  theme_bw()+
  theme(panel.grid.minor = element_blank(), legend.position = "bottom")+
  labs(title = substitute(""^{a}*la*"/"^{b}*la*" of bulk biomass and protein", list(a = a, b = b, la = la)))
```
```{r}
# save the file
ggsave(str_c(outIRMS,"IRMS-hour_allpep_Calisp-colored",".png"), plot = last_plot(), device = "png", width = 10, height = 6)
ggsave(str_c(outIRMS,"IRMS-hour_allpep_Calisp-colored",".svg"), plot = last_plot(), device = "svg", width = 10, height = 6)
```


## 2. Unweighted vs pattern intenstity weighted median ratio as sample protein ratio

Make a long table for plot.

Here only to plot "all good quality peptides"
```{r}
data_calisp_irms_weightlong = data_calisp_irms %>%
  filter(uniq_pep == 0) %>%
  pivot_longer(cols = c(ratio_na, ratio_na.weight),values_to = "ratio_calisp", names_to = "ratio_calisp_weight") %>%
  mutate(ratio_calisp_weight = ifelse(ratio_calisp_weight == "ratio_na", "unweighted", "pattern intensity weighted"))
```

```{r}
data_calisp_irms_weightlong %>%
  ggplot(aes(y = ratio_irms, shape = exp))+
  geom_point(aes(x= ratio_calisp, color = ratio_calisp_weight),size =5.4, alpha = 0.8) +
  geom_text(aes(x= ratio_calisp,label=time), color="White", size = 3.8)+
  xlab(substitute("Protein "^{a}*la*"/"^{b}*la*" (Calis-p Protein-SIP)", list(a = a, b = b, la = la)))+
  ylab(substitute("Total Biomass "^{a}*la*"/"^{b}*la*" (CF-EA-IRMS)", list(a = a, b = b, la = la)))+
  theme_bw()+
  theme(panel.grid.minor = element_blank(), legend.title = element_blank(), legend.position = "bottom")
```
```{r}
# save the file
ggsave(str_c(outIRMS,"IRMS-Calisp_(un)weight",".png"), plot = last_plot(), device = "png", width = 10, height = 6)
ggsave(str_c(outIRMS,"IRMS-Calisp_(un)weight",".svg"), plot = last_plot(), device = "svg", width = 10, height = 6)
```

Plot ratio_calisp against hour, with IRMS ratio (ratio_irms) from all good peptides as color gradient
```{r}
if (la == "C") {
  data_calisp_irms_weightlong %>%
  ggplot(aes(x = hours, y = ratio_calisp))+
  geom_point(aes(color = ratio_irms, shape = ratio_calisp_weight),size = 5.4, alpha = 0.8) +
  scale_shape_manual(values = c(15, 16), name="")+
  scale_color_gradient(low= "#5063f2", high = "#f7743b", name = substitute("Total Biomass "^{a}*la*"/"^{b}*la*" (CF-EA-IRMS)", list(a = a, b = b, la = la))) +
  xlab("Incubation Hours")+
  ylab(substitute("Protein "^{a}*la*"/"^{b}*la*" (Calis-p Protein-SIP)", list(a = a, b = b, la = la)))+
  theme_bw()+
  theme(panel.grid.minor = element_blank(), legend.position = "right")+
  labs(title = substitute(""^{a}*la*"/"^{b}*la*" of bulk biomass and protein", list(a = a, b = b, la = la))) +
  # add seperate corrlation lines for unweighted and weighted group
  geom_smooth(data = data_calisp_irms_weightlong %>% filter(ratio_calisp_weight == "unweighted"), method = "lm", se = FALSE, color = "black", linetype = "dashed") +
  geom_smooth(data = data_calisp_irms_weightlong %>% filter(ratio_calisp_weight == "pattern intensity weighted"), method = "lm", se = FALSE, color = "black", linetype = "solid") +
  # label the correlation lines
  annotate("text", x = 35, y = 1.058, label = "unweighted", color = "black", size = 4) +
  annotate("text", x = 35, y = 1.07, label = "pattern intensity weighted", color = "black", size = 4)
} else {
  data_calisp_irms_weightlong %>%
  ggplot(aes(x = hours, y = ratio_calisp))+
  geom_point(aes(color = ratio_irms, shape = ratio_calisp_weight),size = 5.4, alpha = 0.8) +
  scale_shape_manual(values = c(15, 16), name="")+
  scale_color_gradient(low= "#5063f2", high = "#f7743b", name = substitute("Total Biomass "^{a}*la*"/"^{b}*la*" (CF-EA-IRMS)", list(a = a, b = b, la = la))) +
  xlab("Incubation Hours")+
  ylab(substitute("Protein "^{a}*la*"/"^{b}*la*" (Calis-p Protein-SIP)", list(a = a, b = b, la = la)))+
  theme_bw()+
  theme(panel.grid.minor = element_blank(), legend.position = "right")+
  labs(title = substitute(""^{a}*la*"/"^{b}*la*" of bulk biomass and protein", list(a = a, b = b, la = la))) +
  # add seperate corrlation lines for unweighted and weighted group
  geom_smooth(data = data_calisp_irms_weightlong %>% filter(ratio_calisp_weight == "unweighted"), method = "lm", se = FALSE, color = "black", linetype = "dashed") +
  geom_smooth(data = data_calisp_irms_weightlong %>% filter(ratio_calisp_weight == "pattern intensity weighted"), method = "lm", se = FALSE, color = "black", linetype = "solid") +
  labs(caption = "Linear regression for unweighted(dash line) and weighted (solid line)")
}

  
```
```{r}
# save the file
ggsave(str_c(outIRMS,"Calisp-hour_IRMScolor_(un)weight",".png"), plot = last_plot(), device = "png", width = 10, height = 6)
ggsave(str_c(outIRMS,"Calisp-hour_IRMScolor_(un)weight",".svg"), plot = last_plot(), device = "svg", width = 10, height = 6)
```