---
title: "GEL-sip RAW data wrangling"
output:
  html_document:
    df_print: paged
---

R markdown using calisp.py output. Importing and wrangling of feather data adapted form MS python jupyter notebook.

Here we prepare data in two ways in terms of assigning peptides that mapped to a cluster of multiple proteins and bins.

  1. Same filtering criteria as the one used to plot peptide ratio (data_unique): unique_peptides.csv
  2. Keep the peptides that maps to multiple bins, since this does not impact the bulk analysis on sample level (data_all): good_peptides.csv
  

## Load libraries and basic settings

Setup working directory
```{r setup, include=FALSE}
# setup chuck to make working directory set for all chunks
wd = "/yihualiu/mscalisp/"
# setwd(wd)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = wd)
```

Load libraries

```{r}
library("arrow")
library("tidyverse")
library("tidytable")
```


## Input Parameters

First choice from N (15N) and C (13C), specify the taxonomic reference file and IRMS reference file. 

The input feather files for 13C are in "./GEL_C/calisp_out", for 15N are in "./GEL_N/calisp_out"
```{r}
# la = "C"                                              # specify the isotope label, "C" for 13C and "N" for 15N
la = "N"
sodataxo_dir = str_c(wd,"Soda lake MAGs short taxonomy.csv")   # directory of reference file with bin taxonomic annotation
irmsfile = str_c(wd,"2019_IRMS.csv")                  # specify the IRMS file name/directory
```

Then specify the filter applied to remove unwanted spectra (either NA, "fft", or "clumpy"), and threshold if required at "fft" or "clumpy".
```{r}
# Because the labeling is so slight, we can use the fft error to filter out contaminated spectra
spectra_filter = "fft"
spectra_filter_threshold = 0.0025
```

Generate the other variables depends on C or N

```{r}
if (la == "C") {
  # calisp feather files dir
  calisp_data_dir = "GEL_C/calisp_out/"
  # output dir
  outroot = str_c("GEL_C/wrangled_output/")
  dir.create(outroot)
  isolim = 2   # "2%" for C, "5.643%" for N     # specify the isotope ratio cutoff for Calis-p, defined by the isotope ratio of labelling chemicals, as biomass cannot get higher ratio than label.
} else {
  # calisp feather files dir
  calisp_data_dir = "GEL_N/calisp_out/"
  # output dir
  outroot = str_c("GEL_N/wrangled_output/")
  dir.create(outroot)
  isolim = 5.643  
}
```

## Import data and basic wrangling

```{r}
# acquire file list 

data_raw = NULL
for (f in list.files(calisp_data_dir, full.names = TRUE, pattern = "*.feather")) {
  data_raw = bind_rows(data_raw, read_feather(f))}

  # calculate deltas
  # std used came from Calis-p 
  # C VPDB standard 13C/12C = 0.011056585166521 / 0.988943414833479 
  # 15N/14N = 0.366/99.634 (https://ciaaw.org/nitrogen-references.htm#:~:text=Nitrogen%20Isotope%20Materials,of%20zero%20by%20international%20agreement.; https://doi.org/10.1515/pac-2013-1023)
 
if (la == "C") {
  data_raw = data_raw %>%
    mutate(., delta_na = (ratio_na / (0.011056585166521 / 0.988943414833479)-1)*1000,
               delta_fft = (ratio_fft / (0.011056585166521 / 0.988943414833479)-1)*1000)
} else {
  data_raw = data_raw %>%  
    mutate(., delta_na = (ratio_na / (0.00366/0.99634)-1)*1000,
              delta_fft = (ratio_fft / (0.00366/0.99634)-1)*1000)
}

data_raw = data_raw %>%
  # convert ratio to "%" ratio and for psm
  mutate(ratio_fft = ratio_fft * 100,
          ratio_na = ratio_na * 100,
          psm_neutrons = psm_neutrons * 100)

spectra_load = nrow(data_raw)
print(sprintf("Loaded %d spectra from file(s)", nrow(data_raw)))

```

Optional: Fix the bin name, if wrong delimiter was used and all bins was shown as "unbinned" and all flag_peptide_assigned_to_multiple_bins == FALSE
```{r}
data_raw = data_raw %>%
  mutate(bins = str_extract(proteins, "\\w+|"), flag_peptide_assigned_to_multiple_bins = !is.na(str_extract(proteins, " \\w+")), proteins_old = proteins, proteins = str_extract(proteins, "(?<=\\|)\\S+"))

```
Additional: For 15N data, fix a typo in ms_run, from "...AB1..." to "...A61..."
```{r}
if (la == "N") {
  data_raw = data_raw %>%
    mutate(ms_run = str_replace(ms_run, "AB1", "A61"))
}
```

Save a copy of raw data
```{r}
# save a copy of raw data
write_csv(data_raw, str_c(outroot, "calisp_out_raw.csv"))
```

### Wrangling

#### 1. Remove crapped bins, filter out unwanted spectra ()
Load saved raw data
```{r}
# Load from previously saved raw data
data_raw = read_csv(str_c(outroot, "calisp_out_raw.csv"))
spectra_load = nrow(data_raw)
print(sprintf("Loaded %d spectra from file(s)", nrow(data_raw)))
```

```{r}
if (is.na(spectra_filter)) {
  data = data_raw %>%
    filter((!grepl("crap", bins, ignore.case = TRUE)) &
           (flag_peak_at_minus_one_pos == FALSE) &
           (flag_spectrum_is_wobbly == FALSE) &
           (flag_psm_has_low_confidence == FALSE) &
           (flag_psm_is_ambiguous == FALSE) &
           (flag_spectrum_is_contaminated == FALSE))
} else if (spectra_filter == "fft") {
  data = data_raw %>%
    filter((!grepl("crap", bins, ignore.case = TRUE)) &
           (error_fft < spectra_filter_threshold))
} else if (spectra_filter == "clumpy") {
  data = data_raw %>%
    filter((!grepl("crap", bins, ignore.case = TRUE)) &
           (error_clumpy < spectra_filter_threshold))
}

print(sprintf(" %d spectra (%.1f%%) remaining after filters.", nrow(data), nrow(data)/spectra_load*100))

```
Filter out entries that doesn't fit iso limit

```{r}
data = data %>%
  filter(ratio_na < isolim)
print(sprintf(" %d spectra (%.1f%%) remaining after filters.", nrow(data), nrow(data)/spectra_load*100))
```

#### 2. Discard colums that no longer needed
```{r}
data = data %>%
  select(ms_run, bins, proteins, peptide, pattern_total_intensity, ratio_na, ratio_fft, error_fft, delta_na, delta_fft, flag_peptide_assigned_to_multiple_bins)

print(sprintf(" %d colums left, they are:", ncol(data)))
print(names(data))

```

#### 3. Extract experiment info from ms_run

For example:

    - "Run7_EXPA7_B01_1700ng" -> exp = Blank, control = TRUE, time = 0, rep = 1
    - "Run7_EXPA7_C01_1700ngt" -> exp = Carbon, control = FALSE, time = 0, rep = 1
    - "Run8_EXPA7_TI2_1700ng" -> exp = Initial, control = TRUE, time = 0, rep = 2
    - "Run8_EXPA7_A01_1700ng" -> exp = Amo, control = FALSE, time = 0, rep = 1

```{r}
data = data %>%
  mutate(exp = str_extract(ms_run, "(?<=_)[:upper:][I0-9]\\d(?=_)") )%>%
  mutate(time = ifelse(str_sub(exp, 2,2) == "I", as.integer(0), as.integer(str_sub(exp, 2,2))),
          exp = str_sub(exp,1,1)) %>%
  mutate(control = ifelse(exp %in% c("B","T"), TRUE, FALSE)) %>%
  mutate(exp = ifelse(exp == "B", "Blank", 
          ifelse(exp == "C", "Carbon", 
            ifelse(exp == "T", "Initial", 
              ifelse(exp == "N", "N2",
                ifelse(exp == "U", "Urea",
                  ifelse(exp == "A", "Ammo", exp))))))) %>%
  mutate(sample_ID = str_extract(ms_run, "(?<=EXPA7_)\\w{3}(?=_)"))

```

#### 4. Add incubation hours

The sampling collection point (time) to incubation hours (hours) relationship is written in the irms file.

time initial is assigned as 0
```{r}
timetable <- read_csv(irmsfile) %>%
  summarise(.by = c(Time, Hours)) %>%
  transmute(time = as.integer(Time), hour = as.numeric(Hours)) %>%
  na.omit()

data_all <- inner_join(x = data, y = timetable, by = "time")
```
#### 5. Assign peptides that mapped to a cluster of multiple proteins and bins

Here I prepare the Calis-p data in two ways:
  1. Same filtering criteria as the one used to plot peptide ratio (data_unique)
  2. Keep the peptides that maps to multiple bins, since this does not impact the bulk analysis on sample level (data_all)

```{r}
data_unique = data_all %>%
  filter(flag_peptide_assigned_to_multiple_bins == FALSE)

print(sprintf(" %d spectra (%.1f%%) remaining after filters.", nrow(data_unique), nrow(data_unique)/spectra_load*100))
```
#### 6. Add taxomonic annotation to bins in data_unique 
```{r}
# data_unique = read_csv(str_c(outroot, "calisp_out_unique_peptides.csv"))
# load taxo reference
sodataxo = read_csv(sodataxo_dir)

# Add short name and taxo info to bins, unbinned are marked as "unbinned" in taxo
data_taxo <- sodataxo %>%
  select(bins = bin_ID, phylum, short_name, taxonomy) %>%
  left_join(data_unique,., by = "bins") %>%
  mutate(taxonomy = ifelse(grepl("unbinned", bins), "unbinned", taxonomy),
          phylum = ifelse(grepl("unbinned", bins), "unbinned", phylum),
          short_name = ifelse(grepl("unbinned", bins), "unbinned", short_name))
print(sprintf(" %d spectra (%.1f%%) remaining after filters.", nrow(data_taxo), nrow(data_taxo)/spectra_load*100))
```

#### Save a copy of wrangled data
```{r}
write_csv(data_all, str_c(outroot, "calisp_out_good_peptides.csv"))
write_csv(data_unique, str_c(outroot, "calisp_out_unique_peptides.csv"))
write_csv(data_taxo, str_c(outroot, "calisp_out_unique_peptides_bin_taxo.csv"))
```
