---
title: "Community composition and peptide ratio distribution"
output:
  html_document: default
  pdf_document: default
---

Part 2 of the GEL Calis-p sip data analysis. R markdown using wrangeled calisp.py output. Wrangeling by upstream "GEL_sip_anal_1_raw_wrangling_yl.Rmd" gathers raw binary ".feather" data, cleaned the table and exported to a summerised ".csv". 

Aims of this second part of the analysis is to investigate the data and plot the C isotope distribution of peptides over time

## Load libraries and basic settings

Setup working directory
```{r setup, include=FALSE}
# setup chuck to make working directory set for all chunks
wd = "/yihualiu/mscalisp/"
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = wd)
```

Load libraries

```{r}
library("tidyverse")
library("stringr")
library("ggplot2")
library("tidyr")
library("dplyr")
library("readr")
library("forcats")
library("tidytable")
library("RColorBrewer")
library("Hmisc")
```


## Input Parameters
First choice from N (15N) and C (13C), and also load taxo ref

The input wrangeled files for 13C are in "./GEL_C/wrangeled_output", for 15N are in "./GEL_N/wrangeled_output"
```{r}
la = "C"                                              # specify the isotope label, "C" for 13C and "N" for 15N # nolint
# la = "N"
binls = c("bin25_GEM2", "bin31_GEM2", "bin17_LCM1")   # specify the bins to address when visualizing peptide isotope ratio distribution # nolint
sampsub = "Cyanobacteriota"                           # specify the GTDB PHYLUM to highlight as a subgroup when reporting the isotope ratio of each sample.(Will be reported as the isootpe ratio of selected phylum, other bins, and unbinned) # nolint

sodataxo_dir = str_c(wd,"Soda lake MAGs short taxonomy.csv")   # directory of reference file with bin taxonomic annotation
# irmsfile = str_c(wd,"2019_IRMS.csv")                  # specify the IRMS file name/directory # nolint
#file.choose()

```

Generate the other variables depends on C or N

```{r}
# wrangeled calisp data dir
  wrangeled_data_dir = str_c("GEL_",la,"/wrangled_output/")
# output dir
  outroot = str_c("GEL_",la,"/analysis_", Sys.Date(),"/")
  dir.create(outroot)
```

## Import dataset

```{r}
data_taxo = read_csv(str_c(wrangeled_data_dir,"calisp_out_unique_peptides_bin_taxo.csv"))
# sodataxo = read_csv(sodataxo_dir)
```


## Community composition
Create an output directory for community composition analysis
```{r} 
dir.create(str_c(outroot,"community_composition"))
out_compo = str_c(outroot,"community_composition/")
```

### summarise community composition  
For each samples
```{r}
community_structure_all = data_taxo %>%
  summarise(spectra_count = n(), sum_pattern_intensity = sum(pattern_total_intensity), .by = c(sample_ID ,bins, short_name, time, phylum, taxonomy)) %>%
  arrange(.,desc(spectra_count))%>%
  mutate(MAG_taxo = paste0(short_name,"-", taxonomy))
write_csv(community_structure_all,str_c(out_compo,"community_structure_all_samples.csv"))
community_structure_all
```
Summarized across samples
```{r}
community_structure_sum = data_taxo %>%
  summarise(spectra_count = n(), sum_pattern_intensity = sum(pattern_total_intensity), .by = c(bins, short_name, time, phylum, taxonomy)) %>%
  arrange(.,desc(spectra_count))%>%
  mutate(MAG_taxo = paste0(short_name,"-", taxonomy))
write_csv(community_structure_all,str_c(out_compo,"community_structure_sum.csv"))
community_structure_sum
```

Top 10 bins across all samples (do not include unbinned)  
Note: this is the 10 most abundant bins 
```{r}
t10bins = community_structure_sum %>%
  summarise(spectra_count = sum(spectra_count), sum_pattern_intensity = sum(sum_pattern_intensity), .by = c("bins","short_name","phylum","taxonomy","MAG_taxo")) %>%
  filter(short_name != "unbinned") %>%
  arrange(desc(spectra_count)) %>%
  slice(1:10) %>%
  pull(bins)
t10bins
```

### Plot calis-p based community structure
Plot from T0 and TI
#### Bar plots
**Structure of each samples**
1. With sum_pattern_intensity as abundance
```{r}
# y = sum_pattern_intensity
community_structure_all %>%
  filter(bins %in% t10bins) %>%
  ggplot(aes(sample_ID, sum_pattern_intensity, fill = MAG_taxo))+
    geom_bar(stat = "identity" ,position = "fill")+
    xlab("Sample ID")+
    ylab("Sum of pattern intensity")+
    scale_fill_brewer(palette = "Paired")+
    # scale_fill_manual(values = brewer.pal(10,"Set3"))
    # add title
    labs(title = str_c("Community structure, peptide pattern intensity as MAG abundance, ",la,"-labelled samples"))+
    theme_minimal()+
    theme(plot.title = element_text(size = 12), plot.title.position = "plot", axis.text.x = element_text(angle = 45),
          panel.grid = element_blank())

ggsave(str_c(out_compo,"BarPlot_all_samples_sum_pattern_intensity.svg"), device = svg, width = 10, height = 5, units = "in")
ggsave(str_c(out_compo,"BarPlot_all_samples_sum_pattern_intensity.png"), device = png, width = 10, height = 5, units = "in")
```
2. With spectra_count as abundance
```{r}
# y = spectra_count
community_structure_all %>%
  filter(bins %in% t10bins) %>%
  ggplot(aes(sample_ID, spectra_count, fill = MAG_taxo))+
    geom_bar(stat = "identity" ,position = "fill")+
    xlab("Sample ID")+
    ylab("Number of spectray")+
    scale_fill_brewer(palette = "Paired")+
    labs(title = str_c("Community structure, MS spectra count as MAG abundance, ",la,"-labelled samples"))+
    theme_minimal()+
    theme(plot.title = element_text(size = 12), plot.title.position = "plot", axis.text.x = element_text(angle = 45),
          panel.grid = element_blank())
ggsave(str_c(out_compo,"BarPlot_all_samples_spectra_count.svg"), device = svg, width = 10, height = 5, units = "in")
ggsave(str_c(out_compo,"BarPlot_all_samples_spectra_count.png"), device = png, width = 10, height = 5, units = "in")
```
**Community structure of T0 samples (time = 0)**
1. With sum_pattern_intensity as abundance
```{r} 
# y = sum_pattern_intensity
community_structure_all %>%
  filter(time == 0 & bins %in% t10bins) %>%
  ggplot(aes(sample_ID, sum_pattern_intensity, fill = MAG_taxo))+
    geom_bar(stat = "identity" ,position = "fill")+
    xlab("Sample ID")+
    ylab("Sum of pattern intensity")+
    scale_fill_brewer(palette = "Paired")+
    labs(title = str_c("Community structure, peptide pattern intensity as MAG abundance, ",la,"-T0 & blank samples"))+
    theme_minimal()+
    theme(plot.title = element_text(size = 12), plot.title.position = "plot", axis.text.x = element_text(angle = 45),
          panel.grid = element_blank())
ggsave(str_c(out_compo,"BarPlot_T0_samples_sum_pattern_intensity.svg"), device = svg, width = 10, height = 5, units = "in")
ggsave(str_c(out_compo,"BarPlot_T0_samples_sum_pattern_intensity.png"), device = png, width = 10, height = 5, units = "in")
```

2. With sum_pattern_intensity as spectra_count
```{r}
# y = spectra_count
community_structure_all %>%
  filter(time == 0 & bins %in% t10bins) %>%
  ggplot(aes(sample_ID, spectra_count, fill = MAG_taxo))+
    geom_bar(stat = "identity" ,position = "fill") +
    xlab("Sample ID") +
    ylab("Number of spectra")+
    scale_fill_brewer(palette = "Paired")+
    labs(title = str_c("Community structure, MS spectra count as MAG abundance, ",la,"-T0 & blank samples"))+
    theme_minimal()+
    theme(plot.title = element_text(size = 12), plot.title.position = "plot", axis.text.x = element_text(angle = 45),
          panel.grid = element_blank())
ggsave(str_c(out_compo,"BarPlot_T0_samples_spectra_count.svg"), device = svg, width = 10, height = 5, units = "in")
ggsave(str_c(out_compo,"BarPlot_T0_samples_spectra_count.png"), device = png, width = 10, height = 5, units = "in")
```

**structure of the summarized community composition**
```{r}
community_structure_sum %>%
  filter(bins %in% t10bins) %>%
  pivot_longer(cols = c(sum_pattern_intensity, spectra_count), names_to = "MAG_abundance_from", values_to = "relative_abundance") %>%
  ggplot(aes(MAG_abundance_from, relative_abundance , fill = MAG_taxo))+
    geom_bar(stat = "identity" ,position = "fill")+
    scale_fill_brewer(palette = "Paired")+
    labs(title = str_c("Overall community structure across all ",la," labeled & blank samples"))+
    theme_minimal()+
    theme(plot.title = element_text(size = 12), plot.title.position = "plot",
          panel.grid = element_blank())
ggsave(str_c(out_compo,"BarPlot_summarised_samples.svg"), device = svg, width = 10, height = 5, units = "in")
ggsave(str_c(out_compo,"BarPlot_summarised_samples.png"), device = png, width = 10, height = 5, units = "in")
```

**community structure at phylum level**
```{r}
community_structure_sum %>%
  pivot_longer(cols = c(sum_pattern_intensity, spectra_count), names_to = "means_of_sum", values_to = "relative_abundance") %>%
  ggplot(aes(means_of_sum, relative_abundance , fill = phylum))+
  geom_bar(stat = "identity" ,position = "fill")+
  labs(title = str_c("Phylum level community structure across all ",la," labeled & blank samples"))+
  theme_minimal()+
  theme(plot.title = element_text(size = 12), plot.title.position = "plot",
          panel.grid = element_blank())
ggsave(str_c(out_compo,"BarPlot_summarised_samples_all_bin_phylum.svg"), device = svg, width = 10, height = 5, units = "in")
ggsave(str_c(out_compo,"BarPlot_summarised_samples_all_bin_phylum.png"), device = png, width = 10, height = 5, units = "in")
```


#### Bubble plots

Make bubble plot showing the relative abundance of most abundant bins in T0 samples. 

**Prepare data**
```{r}
bubble_data <- community_structure_all %>%
  filter(time == 0, short_name != "unbinned") %>%
  # Normalise the sum_pattern_intensity among all bins in each sample
  mutate(sum_pattern_intensity_norm = sum_pattern_intensity/sum(sum_pattern_intensity),
          spectra_count_norm = spectra_count/sum(spectra_count), .by = "sample_ID")

# number of samples
n_samples <- bubble_data %>%
  pull(sample_ID) %>%
  unique() %>%
  length()

# Filter out bins that has relative abundance >= 0.001.
# 1. With normalised sum_pattern_intensity as relative abundance
  bubble_data_SI <- bubble_data %>%
  filter(sum_pattern_intensity_norm >= 0.001)
  # Keep only bins that still present in all samples after filtering.
  bin_occurence <- bubble_data_SI %>%
    group_by(MAG_taxo) %>%
    summarise(occurence = sum(sum_pattern_intensity_norm > 0)) %>%
    filter(occurence == n_samples) %>%
    pull(MAG_taxo)
  bubble_data_SI <- bubble_data_SI %>%
    filter(MAG_taxo %in% bin_occurence)%>%
    # Reorder the MAG_taxo to plot on y axis from top to bottom alphabetically.
    mutate(MAG_taxo = factor(MAG_taxo, levels = unique(MAG_taxo)[order(unique(MAG_taxo), decreasing = TRUE)]))
# 2. With normalised spectra_count as relative abundance
  bubble_data_SC <- bubble_data %>%
  filter(spectra_count_norm >= 0.001)
  # Keep only bins that still present in all samples after filtering.
  bin_occurence <- bubble_data_SC %>%
    group_by(MAG_taxo) %>%
    summarise(occurence = sum(spectra_count_norm > 0)) %>%
    filter(occurence == n_samples) %>%
    pull(MAG_taxo)
  bubble_data_SC <- bubble_data_SC %>%
    filter(MAG_taxo %in% bin_occurence)%>%
    # Reorder the MAG_taxo to plot on y axis from top to bottom alphabetically.
    mutate(MAG_taxo = factor(MAG_taxo, levels = unique(MAG_taxo)[order(unique(MAG_taxo), decreasing = TRUE)]))

# save the data
write_csv(bubble_data, str_c(out_compo, "bubble_plot_data.csv"))
write_csv(bubble_data_SI, str_c(out_compo, "bubble_plot_data_norm_intensity_filtered.csv"))
write_csv(bubble_data_SC, str_c(out_compo, "bubble_plot_data_norm_count_filtered.csv"))
```

**Make bubble plot**
```{r}
# Plot with normalised sum_pattern_intensity as relative abundance, colour the bubbles by phylum
bubble_data_SI %>%
  ggplot(aes(x = sample_ID, y = MAG_taxo)) +
  geom_count(aes(size = sum_pattern_intensity_norm, colour = phylum), alpha = 0.8)+
    scale_size(range = c(0,15))+
    guides(size = guide_legend("Relative abundance", size = c(1,2,5)))+
    theme_classic() + 
    xlab("Sample ID")+
    ylab("MAGs")+
    scale_color_brewer(palette = "Set1")
# save the plot
ggsave(str_c(out_compo,"BubblePlot_pattern_intensity_as_abundance.svg"), device = svg, width = 6, height = 5, units = "in")
ggsave(str_c(out_compo,"BubblePlot_pattern_intensity_as_abundance.png"), device = png, width = 6, height = 5, units = "in")
```

```{r}
# Plot with normalised spectra_count as relative abundance, colour the bubbles by phylum
bubble_data_SC %>%
  ggplot(aes(x = sample_ID, y = MAG_taxo)) +
  geom_count(aes(size = spectra_count_norm, colour = phylum), alpha = 0.8)+
    scale_size(range = c(0,15))+
    guides(size = guide_legend("Relative abundance", size = c(1,2,5)))+
    theme_classic() + 
    xlab("Sample ID")+
    ylab("MAGs")+
    scale_color_brewer(palette = "Set1")
# save the plot
ggsave(str_c(out_compo,"BubblePlot_spectra_count_as_abundance.svg"), device = svg, width = 6, height = 5, units = "in")
ggsave(str_c(out_compo,"BubblePlot_spectra_count_as_abundance.png"), device = png, width = 6, height = 5, units = "in")
```


## Peptide distribution at each time point (15/14N or 13C/12C ~ timepoint)
### 1. peptide 13C/12C or 15N/14N
Here we plot the neutron-abundance-model-estimated 13C/12C or 15N/14N for the three top populations as a function of time.

Prepare dataset
```{r}
# Prepare dataset for peptide distribution box plots

# create a directory for peptide distribution
boxdir = str_c(outroot, "/peptide_distribution/")
dir.create(boxdir)

# Function to perform Wilcoxon test between a group and each control
perform_wilcox_test <- function(exp_data, control_data) {
  test_result <- wilcox.test(exp_data$ratio_na, control_data$ratio_na)
  return(test_result$p.value)
}

# Function to perform Kruskal-Wallis test on the specified groups
perform_kruskal_test <- function(all_data) {
  test_result <- kruskal.test(ratio_na ~ control, data = all_data)
  return(test_result$p.value)
}

# Prepare to perform statistical tests using the original data_taxo
test_res <- data.frame()

# Loop through each unique bin/taxonomy combination
for (bin in unique(data_taxo$bins)) {
  for (time in unique(data_taxo$time)) {
    # Subsetting data from the original data_taxo
    blank_data <- data_taxo %>% filter(bins == bin, time == time, exp == "Blank")
    initial_data <- data_taxo %>% filter(bins == bin, time == 0, exp == "Initial")
    combined_control_data <- data_taxo %>% filter(bins == bin, control == TRUE)
    
    for (exp_name in unique(data_taxo$exp)){
      exp_data <- data_taxo %>% filter(bins == bin, time == time, exp == exp_name)
    
    # Initialize p-values
    p_value_blank <- NA
    p_value_initial <- NA
    p_value_all_control <- NA

    # Wilcox test vs Blank
    if (nrow(blank_data) > 0 & nrow(exp_data) > 0) {
      p_value_blank <- perform_wilcox_test(select(exp_data,ratio_na), select(blank_data,ratio_na))
    }

    # Wilcox test vs Initial
    if (nrow(initial_data) > 0 & nrow(exp_data) > 0) {
      p_value_initial <- perform_wilcox_test(select(exp_data,ratio_na), select(initial_data,ratio_na))
    }

    # Wilcox test vs all control (Blank + Initial)
    if (nrow(combined_control_data) > 0 & nrow(exp_data) > 0) {
      p_value_all_control <- perform_wilcox_test(select(exp_data,ratio_na), select(combined_control_data,ratio_na))
    }

    # Add results to test_res
    test_res <- bind_rows(test_res, tibble(bins = bin,
                                            exp = exp_name,
                                            time = time, 
                                            taxonomy = unique(exp_data$taxonomy),
                                            control = FALSE,
                                            exp_vs_blk_p_value = p_value_blank,
                                            exp_vs_init_p_value = p_value_initial,
                                            exp_vs_all_control_p_value = p_value_all_control))
  
    }  
  }
}

# Add adjust p-values
test_res_padj <- test_res %>%
  mutate(adj_exp_vs_blk_p_value = p.adjust(exp_vs_blk_p_value, method = "BH"),
         adj_exp_vs_init_p_value = p.adjust(exp_vs_init_p_value, method = "BH"),
         adj_exp_vs_all_control_p_value = p.adjust(exp_vs_all_control_p_value, method = "BH"))
```


```{r}

# Filter out the selected bins of interest and calculate values to plot
weighted_box_stats = data_taxo %>%
  select(bins, sample_ID, exp, time, hour, control, pattern_total_intensity, ratio_na, taxonomy) %>%
  group_by(bins, exp, time,hour, taxonomy, control) %>%
  summarise(minRo = min(ratio_na),
            y25 = wtd.quantile(ratio_na, weights = pattern_total_intensity, probs = 0.25),
            y50 = wtd.quantile(ratio_na, weights = pattern_total_intensity, probs = 0.5),
            y75 = wtd.quantile(ratio_na, weights = pattern_total_intensity, probs = 0.75),
            maxRo = max(ratio_na)) %>%
  # Note: in geom_boxplot, ymin is not the min number of y, but  
  #  "smallest observation greater than or equal to lower hinge (q1) - 1.5 * IQR"
  # where IQR ( inter-quartile range) = q3-q1
  # here instead, the following code use min of y or q1 - 1.5*IQR, whichever greater
  # similarly, ymax = max of y or q3 + 1.5*IQR, whichever smaller
  mutate(yIQR = y75 - y25) %>%
  mutate(y0 = ifelse(y25 - 1.5 * yIQR > minRo, y25 - 1.5 * yIQR, minRo),
         y100 = ifelse(y75 + 1.5 * yIQR < maxRo, y75 + 1.5 * yIQR, maxRo)) %>%
  # join the quantile information with the stastics
  left_join(., test_res_padj, by = c("bins", "exp", "time","taxonomy","control"))


```

```{r}
unweighted_box_stats = data_taxo %>%
  select(bins, sample_ID, exp, time, hour, control, ratio_na, taxonomy) %>%
  group_by(bins, exp, time,hour, taxonomy, control) %>%
  summarise(minRo = min(ratio_na),
            y25 = wtd.quantile(ratio_na, probs = 0.25),
            y50 = wtd.quantile(ratio_na, probs = 0.5),
            y75 = wtd.quantile(ratio_na, probs = 0.75),
            maxRo = max(ratio_na)) %>%
  # Note: in geom_boxplot, ymin is not the min number of y, but  
  #  "smallest observation greater than or equal to lower hinge (q1) - 1.5 * IQR"
  # where IQR ( inter-quartile range) = q3-q1
  # here instead, the following code use min of y or q1 - 1.5*IQR, whichever greater
  # similarly, ymax = max of y or q3 + 1.5*IQR, whichever smaller
  mutate(yIQR = y75 - y25) %>%
  mutate(y0 = ifelse(y25 - 1.5 * yIQR > minRo, y25 - 1.5 * yIQR, minRo),
         y100 = ifelse(y75 + 1.5 * yIQR < maxRo, y75 + 1.5 * yIQR, maxRo))%>%
  # join the quantile information with the stastics
  left_join(., test_res_padj, by = c("bins", "exp", "time","taxonomy","control"))

# Export boxplot data 
write.csv(weighted_box_stats, str_c(boxdir,la, "_weighted_stats_2019GEL_pro-SIP.csv"))
write.csv(unweighted_box_stats, str_c(boxdir,la, "_unweighted_stats_2019GEL_pro-SIP.csv"))

weighted_box_plot = filter(weighted_box_stats, bins %in% binls) %>%
  # 2024-12-01 assign "Blank", time -1 and hour -1 to initial samples to fit in the same plot
  mutate(hour = if_else(exp == "Initial", -1, hour)) %>%
  mutate(time = if_else(exp == "Initial", -1, time)) %>%
  mutate(exp = if_else(exp == "Initial", "Blank", exp))
weighted_box_plot %>%
write.csv(., str_c(boxdir,la, "_weighted_stats_plotted_2019GEL_pro-SIP.csv"))

unweighted_box_plot = filter(unweighted_box_stats, bins %in% binls) %>%
  mutate(hour = if_else(exp == "Initial", -1, hour)) %>%
  mutate(time = if_else(exp == "Initial", -1, time)) %>%
  mutate(exp = if_else(exp == "Initial", "Blank", exp)) 
unweighted_box_plot %>%
write.csv(., str_c(boxdir,la, "_unweighted_stats_plotted_2019GEL_pro-SIP.csv"))

```

Function to plot

```{r}
#test
# stats = weighted_box_plot
# weight = "IntWt"

#creating individual box plots
indiv_box = function(stats, weight){
  for (i in binls) {
  stats_i <- filter(stats, bins == i) %>%
    mutate(hour = as.numeric(hour),
           exp_color = ifelse(exp == "Blank" & hour == -1 & time == -1, "Initial", exp))  # New color variable
  
  # calculate the blank quantiles' average over time, as reference lines to show on the diagram
  blk <- stats_i %>%
    filter(control == TRUE)
  
  y25ave = mean(blk$y25)
  y50ave = mean(blk$y50)
  y75ave = mean(blk$y75)

  # get taxo annotation
  taxo_i = stats_i[1,]$taxonomy
    
  # plot
  if (la == "C") {
    stats_i %>% 
    ggplot(., aes(hour, group = hour, fill = exp_color)) +
    geom_boxplot(aes(x= hour, 
                     ymin = y0, 
                     lower = y25, middle = y50, upper = y75, ymax = y100 
                     ), alpha = 0.8,
                 stat = "identity")+
    facet_grid(~exp)+
    scale_fill_manual(values = c("Initial" = "#BEBEBE","Blank" = "#59a9eb", "Carbon" = "#25c450"))+
    ylim(0.5, 1.5)+
    ylab( str_c("Peptide 13C/12C (%) distribution"))+
    theme_bw()+
    #Add lines for the mean y75, mean y75 and mean y50 of blank, values were calculated externally
    geom_hline(aes(yintercept = y75ave), linetype = 1, color = "#949494")+
    geom_hline(aes(yintercept = y25ave), linetype = 1, color = "#949494")+
    geom_hline(aes(yintercept = y50ave), linetype = 2, color = "#949494")+
    labs(title = str_c(i," ", taxo_i, " ",weight," ",la," isotope distribution"), fill = "Experiment")
    
  } else {
    stats_i %>% 
      ggplot(., aes(hour, group = hour, fill = exp_color)) +
      geom_boxplot(aes(x= hour, 
                       ymin = y0, 
                       lower = y25, middle = y50, upper = y75, ymax = y100 
                       ), alpha = 0.8,
                   stat = "identity")+
      facet_grid(~exp)+
      scale_fill_manual(values = c("Initial" = "#BEBEBE","Blank" = "#59a9eb", "Carbon" = "#25c450", "Ammo" = "#c2083c", "Urea" = "#ffcd00", "N2" = "#e0c4ad"))+
      ylim(-1.5, 2)+
      ylab( str_c(i, " Protein 15N/14N (%)"))+
      theme_bw()+
      #Add lines for the mean y75, mean y75 and mean y50 of blank, values were calculated externally
      geom_hline(aes(yintercept = y75ave), linetype = 1, color = "#949494")+
      geom_hline(aes(yintercept = y25ave), linetype = 1, color = "#949494")+
      geom_hline(aes(yintercept = y50ave), linetype = 2, color = "#949494")+
    labs(title = str_c(la,"_",weight,"_",i,"_",taxo_i), fill = "Experiment")
  }

  ggsave(str_c(boxdir,la,"_Boxplot_",weight,"_",taxo_i,"_2019GEL.pdf"), plot= last_plot(), device = "pdf", width = 8, height = 8.5, units = "in")
  ggsave(str_c(boxdir,la,"_Boxplot_",weight,"_",taxo_i,"_2019GEL.png"), plot= last_plot(), device = "png", width = 8, height = 8.5, units = "in")

} 
}

  
```

Ploting box plot
```{r}
indiv_box(weighted_box_plot, "Intensity_Weighted")
indiv_box(unweighted_box_plot, "Unweighted")
```
The ploted data are (previously exported to file):
```{r}
filter(weighted_box_stats, bins %in% binls)
filter(unweighted_box_stats, bins %in% binls)
```

### 2. peptide 13C/12C or 15N/14N difference between control and treatment group

What if we calculate the change on a per-peptide basis from a baseline established by the control data...
 This provides a way to look at the most reliable part of the data as well as eliminate intra-peptide noise 
 by comparing the same peptides accross treatments and controls 
 For each peptide at each timepoint in the treatments, we calculate the difference of the neutron-abundance
 -estimated ratio from the base-line ratio in the controls (calculate the difference at each time point)
 
```{r}
# separate control and treatment data, then filter out non-shred peptides
# control data
c_spectra = filter(data_taxo, (control == 1)&(bins %in% binls))

# unique peptides in control
c_peptide_names = c_spectra %>%
  summarise(.by = peptide)

# treatment data
spectra = filter(data_taxo, (control == 0)&(bins %in% binls)) %>%
  # shared peptides
  filter(., peptide %in% c_peptide_names$peptide)

# Original: creating a peptide table with 'bin', 'time', 'peptide', 'c_#spectra', 'c_na_mean', 'c_na_stdev', 'bt_#spectra', 'bt_na_mean', 'bt_na_stdev'
#  Instead, I use "ctrl" for control ("c"), "num" for #, "na-ratio" for "na", and removed "bt" for treatments
#  Also, using medium instead of mean
c_peptide_df <- c_spectra %>%
  group_by(peptide, bins, time, hour, short_name, taxonomy) %>%
  summarise(., ctrl_spectra_num = n(),
            ctrl_na_ratio_mid = median(ratio_na),
            ctrl_na_ratio_stdev = sd(ratio_na))

peptide_df <- spectra %>%
  select(peptide, bins, ratio_na, time, hour, short_name, taxonomy) %>%
  group_by(peptide, bins, time, hour, short_name, taxonomy) %>%
  summarise(., spectra_num = n(),
            na_ratio_mid = median(ratio_na),
            na_ratio_stdev = sd(ratio_na)) %>%
  inner_join(., c_peptide_df, by = c("peptide", "bins", "time", "hour", "short_name", "taxonomy")) %>%
  drop_na() %>%
  mutate(na_ratio_diff = na_ratio_mid - ctrl_na_ratio_mid,
          bintaxo = str_c(short_name, "_", taxonomy))  
print(nrow(peptide_df))
 
```
Let's plot out the ratio difference of each peptide
```{r}
peptide_df %>%
  ggplot() +
  geom_boxplot(aes(x = hour, y = na_ratio_diff), outlier.shape = NA) +
  geom_hline(yintercept = 0, color = "red", linetype = 5) +
  facet_wrap(~ bintaxo) +
  ylim(-0.35, 0.35) + 
  theme_bw()+
  theme(panel.grid.major = element_line(colour = NA),
    panel.grid.minor = element_line(colour = NA),
    panel.background = element_rect(fill = NA))
```
```{r}
 ggsave(str_c(boxdir,la,"_na_ratio_diff__2019GEL.pdf"), plot= last_plot(), device = "pdf", width = 8, height = 6, units = "in")
  ggsave(str_c(boxdir,la,"_na_ratio_diff__2019GEL.png"), plot= last_plot(), device = "png", width = 8, height = 6, units = "in")
```


